### 2018-05-18 3_encod.tex
add reference to Unfoldings

\section{Portability analysis as a bounded reachability problem}
\label{ch:enc:enc}

@@

\subsection{Encoding for the control-flow} %Encoding the control-flow constraints}
\label{ch:enc:enc:cf}

@@

While working on \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
The changes are conditioned by the need to be able to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
For that, we compile the \textit{Abstract Syntax Tree~(AST)} of the parsed C-code to the plain event-flow graph.
We show %TODO! to show, actually!!!
that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner. This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions. 
For instance, \porthos{} used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
and control-flow of the branching instruction $i_1 = (c\,?\,i_2\!:\!i_3)$ was encoded as
$\phi_{CF}(c\,?\,i_2\!:\!i_3) \defeq (\textit{cf}_{i_1} \Leftrightarrow (\textit{cf}_{i_2} \lor \textit{cf}_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
In contrast, the new scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula. The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:enc:enc:df}).

All control-flow edges contain a label that denotes the \textit{guard}, a predicate that determines the transition.
An empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
A guard depends on the data-flow of the program, it represents a memory unit or a computational expression that is liable to the weak memory model relaxations.
The branching expressions that support more than two outgoing control-flow edges may be useful for describing non-deterministic transition systems, where the guards are not necessarily mutually exclusive.
However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition).

@@

\subsection{Encoding for the data-flow}
\label{ch:enc:enc:df}

@@

\subsection{Encoding for the memory model} %Encoding the memory model constraints}
\label{ch:enc:enc:wmm}

>>>>

\section{Portability analysis as a bounded reachability problem}
\label{ch:enc:bmc}

@@

\subsection{Encoding for the control-flow} %Encoding the control-flow constraints}
\label{ch:enc:bmc:cf}

@@

While working on \porthos[2], we have applied some modifications to the encoding scheme for the control-flow.
These changes were motivated by the need to process an arbitrary control-flow produced by conditional and unconditional jumps of the C language.
For that, we compile the \textit{Abstract Syntax Tree~(AST)} of the parsed C-code to the plain event-flow graph.
We show %TODO! to show, actually!!!
that the new encoding is smaller than the old one used in \porthos{} since it does not produces new variables for each high-level statement of the input language.
%The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner. This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions.

For instance, \porthos[1] used the encoding scheme where the control-flow of the sequential instruction $i_1 = i_2; i_3$ was encoded as
$\phi_{CF}(i_2;i_3) \defeq (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$,
and control-flow of the branching instruction $i_1 = (c\,?\,i_2\!:\!i_3)$ was encoded as
$\phi_{CF}(c\,?\,i_2\!:\!i_3) \defeq (\textit{cf}_{i_1} \Leftrightarrow (\textit{cf}_{i_2} \lor \textit{cf}_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$
(here we used the notation of C-like ternary operator `\texttt{x?y:z}' for defining the conditional expression `$\texttt{if}\;x\;\texttt{then}\;y\;\texttt{else}\;z$').
In contrast, the new encoding scheme implemented in \porthos[2] firstly compiles the recursive high-level code into the linear low-level event-based representation, that is then encoded into an SMT-formula. The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see Section~\ref{ch:enc:bmc:df}).
In general, the new encoding scheme follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} for encoding Petri-nets.

All control-flow edges contain a label that denotes the \textit{guard}, a predicate that determines the transition.
An empty guard (a \textit{true} predicate) is denoted as $\epsilon$.
A guard depends on the data-flow of the program, it represents a memory unit or a computational expression that is liable to the weak memory model relaxations.
The branching expressions that support more than two outgoing control-flow edges may be useful for describing non-deterministic transition systems, where the guards are not necessarily mutually exclusive.
However, as the C language supports only binary logic (\textit{if-then-else} branching), \porthos[2] builds only two possible outcomes of evaluating a guard (\textit{primary} and \textit{alternative} transition), see Section~\ref{ch:impl:model:xgraph:edges} for details.

@@

\subsection{Encoding for the data-flow}
\label{ch:enc:bmc:df}

@@

\subsection{Encoding for the memory model} %Encoding the memory model constraints}
\label{ch:enc:bmc:wmm}

### 2018-05-18 4_impl.tex
cleanup a bit
add reference to Unfoldings

\section{Architecture}
\label{ch:impl:arch}

@@

\begin{figure}%[!h]%[!hb]
\centering
%basicstyle=\ttfamily\tiny,
\begin{lstlisting}[basicstyle=\rmfamily\scriptsize,%
  %alsoletter={:,|,;,(,)},
  %morekeywords={:,|,;,(,)},
  %keywordstyle=\color{dkblue},
  literate=%
  {<}{{$\langle$}}1
  {>}{{$\rangle$}}1,
  escapeinside={[*}{*]},]
<program>
      :      <initialisation>   <thread>+   <assertion>
      ;
<thread>   
      :      thread   <thread-id>   <instruction>
      ;
<instruction>
      :      <atom>
      |      `{'   <instruction>   `}'
      |      <instruction>   `;'   <instruction>
      |      `while'   `('   <bool-expr>   `)'   <instruction>
      |      `if'   <bool-expr>   `{'   <instruction>   `}'   <instruction>
      ;
<atom>   
      :      <register>   `[*<*]-'   <expression>
      |      <register>   `[*<*]:-'   <location>
      |      <location>   `:='   <register>
      |      <register>   `='   <location>   `.'   `load'      `('   <atomic>   `)'
      |      <location>   `='   <register>   `.'   `store'   `('   <atomic>   `)'
      |      (`mfence'   |   `sync'   |   `lwsync'   |   `isync')
      ;
<bool-expr>
      :      `true'
      |      `false'
      |      <expression>   (`and'      |      `or')   <expression>
      |      <expression>   (`=='   |   `!='   |   `>'   |   `>='   |   `[*<*]'   |   `[*>*]=')   <expression>
      ;
<expression>   
      :      [0-9]
      |      <register>
      |      <expression>   (`*'   |   `+'   |   `-'   |   `/'   |   `%')   <expression>
      ;
\end{lstlisting}
\caption{The sketch of the input language grammar used by \oldporthos}
\label{fig:in_grammar_pts}
\end{figure}

@@

\subsection{The internal representations}
\label{ch:impl:model}

@@

\textbf{TODO: say sth about old event hierarchy: very restricted (e.g., only loc<-reg)}

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po{} and the data-flow relations \co{} and \rf{}.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
%The \xgraph{} is constructed by the mutable graph builder. %TODO: name of builder -- after refactoring

@@

\paragraph{Edges}
\label{ch:impl:model:xgraph:edges}
Edges are stored as an immutable hash-map from source to target event.
The type of edge is stored in the enumeration \texttt{XEdgeKind}:%TODO: IMPLEMENT THIS
\begin{itemize}[noitemsep]
  \item the \textit{control-flow edges}: a primary edge (epsilon or if-then transition) and an alternative edge (if-else transition), see more in Section~\ref{ch:port:enc:cf}; and
  \item the \textit{data-flow edges}: the \co{}-relation edge and the \rf{}-relation edge. %TODO! implement and then describe more!
\end{itemize}

\paragraph{Graph invariants}
\label{ch:impl:model:xgraph:invariants}
Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:port:enc:cf}):

@@

\subsection{The processing units} %Transformation}
\label{ch:impl:proc}

>>>>

\section{The architecture}
\label{ch:impl:arch}

@@

\begin{figure}[t]%[!h]%[!hb]
\centering
%basicstyle=\ttfamily\tiny,
\begin{lstlisting}[basicstyle=\rmfamily\scriptsize,%
  %alsoletter={:,|,;,(,)},
  %morekeywords={:,|,;,(,)},
  %keywordstyle=\color{dkblue},
  literate=%
  {<}{{$\langle$}}1
  {>}{{$\rangle$}}1,
  escapeinside={[*}{*]},]
<program>
      :      <initialisation>   <thread>+   <assertion>
      ;
<thread>   
      :      thread   <thread-id>   <instruction>
      ;
<instruction>
      :      <atom>
      |      `{'   <instruction>   `}'
      |      <instruction>   `;'   <instruction>
      |      `while'   `('   <bool-expr>   `)'   <instruction>
      |      `if'   <bool-expr>   `{'   <instruction>   `}'   <instruction>
      ;
<atom>   
      :      <register>   `[*<*]-'   <expression>
      |      <register>   `[*<*]:-'   <location>
      |      <location>   `:='   <register>
      |      <register>   `='   <location>   `.'   `load'      `('   <atomic>   `)'
      |      <location>   `='   <register>   `.'   `store'   `('   <atomic>   `)'
      |      (`mfence'   |   `sync'   |   `lwsync'   |   `isync')
      ;
<bool-expr>
      :      `true'
      |      `false'
      |      <expression>   (`and'      |      `or')   <expression>
      |      <expression>   (`=='   |   `!='   |   `>'   |   `>='   |   `[*<*]'   |   `[*>*]=')   <expression>
      ;
<expression>   
      :      [0-9]
      |      <register>
      |      <expression>   (`*'   |   `+'   |   `-'   |   `/'   |   `%')   <expression>
      ;
\end{lstlisting}
\caption{The sketch of the input language grammar used by \oldporthos}
\label{fig:in_grammar_pts}
\end{figure}

@@

\subsection{Internal representations}
\label{ch:impl:model}

@@

<TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO!!!!!!!!!!!!!!!!!!!!!!>
%TODOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOOO
\textit{\textbf{TODO: say sth about old event hierarchy: very restricted (e.g., only loc<-reg)}}

The \ytree{} is compiled into the low-level event-based program representation called \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po{} and the data-flow relations \co{} and \rf{}.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF].
The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \; \cup \; \xgraph[DF]$.
The UML diagram on Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the hierarchy of main interfaces of the X-abstraction level.
Internally, the graph is represented by an adjacency matrix (to be exact, by multiple adjacency matrices that store edges of different kinds, see more details in Section~\ref{ch:impl:model:xgraph:edges}).

@@

\paragraph{Edges}
\label{ch:impl:model:xgraph:edges}
As the graph is represented by an adjacency matrix, its edges are stored in (immutable) hash-maps.
We distinguish the following kinds of edges:

\begin{outline}%[noitemsep]
  \1 the \textit{control-flow edges}:
    \2 the \textit{primary edges}, that denote both $\epsilon$-labelled transitions (in case of linear sequence of events) and conditional transition that evaluates the conditional event (the source of the transition) to the \textit{true}, and
    \2 the \textit{alternative edges}, that denote conditional transitions for which the conditional event (the source) was evaluated to the \textit{false}; and
  \1 the \textit{data-flow edges}:
    \2 the \co{}-relation edges, and 
    \2 the \rf{}-relation edges. %TODO! implement 
\end{outline}

\paragraph{Graph invariants}
\label{ch:impl:model:xgraph:invariants}
Once being constructed, the graph must conform the following requirements (see also Section~\ref{ch:enc:bmc:cf}):

@@

\subsection{Processing units} %Transformation}
\label{ch:impl:proc}
